---
title: "Segmentation and Maximizing Profit with Supervised Learning Methods"
author: "Fransesca , Valentin Barthel, Adrien Busché"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Packages

```{r carse, echo=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(kohonen)
library(caret)
library(randomForest)
library(pROC)
library(glmnet)
library(Hmisc)

# Load the data
data <- read.csv("marketing_campaign.csv", sep = ";")

# Clean the data
data_clean <- data %>% filter(complete.cases(.))

# Convert Dt_Customer to a numeric feature (days since first acquisition)
data_clean$Dt_Customer <- as.Date(data_clean$Dt_Customer, format = "%Y-%m-%d")
data_clean$Days_Customer <- as.numeric(difftime(data_clean$Dt_Customer, min(data_clean$Dt_Customer), units = "days"))
data_clean <- data_clean %>% select(-Dt_Customer)

# Convert categorical variables
data_clean$Education <- as.factor(data_clean$Education)
data_clean$Marital_Status <- as.factor(data_clean$Marital_Status)
data_clean$Response <- as.factor(data_clean$Response)
```

# Exploratory Data Analysis

## Descriptive Statistics

```{r pressure, echo=FALSE}
summary(data_clean)

```

## Visualization of Response Distribution

```{r pressure, echo=FALSE}
ggplot(data_clean, aes(x = Response)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  labs(title = "Response Distribution", x = "Response", y = "Count")

```

## Segmentation with Self-Organizing Maps (SOM)

```{r pressure, echo=FALSE}
# Normalize numeric data for SOM
numeric_data <- data_clean %>%
  select(Age = Year_Birth, Income, MntWines:MntGoldProds, NumDealsPurchases:NumWebVisitsMonth) %>%
  scale()

# Create SOM grid and train the model
som_grid <- somgrid(xdim = 5, ydim = 5, topo = "hexagonal")
som_model <- som(numeric_data, grid = som_grid, rlen = 100, alpha = c(0.05, 0.01))

# Visualize SOM clusters
plot(som_model, type = "codes")

# Add clusters to the dataset
som_clusters <- cutree(hclust(dist(som_model$codes[[1]])), k = 5)
data_clean$Cluster <- as.factor(som_clusters[som_model$unit.classif])

```

# Model Training and Threshold Optimization

## Data splitting

```{r pressure, echo=FALSE}
set.seed(123)
train_index <- createDataPartition(data_clean$Response, p = 0.7, list = FALSE)
train_data <- data_clean[train_index, ]
test_data <- data_clean[-train_index, ]

```

## Logistic Regression with LASSO Regularization

```{r pressure, echo=FALSE}
# Prepare data for glmnet
x_train <- model.matrix(Response ~ ., data = train_data)[, -1]
y_train <- as.numeric(train_data$Response) - 1
x_test <- model.matrix(Response ~ ., data = test_data)[, -1]

# Train with LASSO
lognet_model <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1)
lognet_pred <- predict(lognet_model, newx = x_test, type = "response", s = "lambda.min")
```

## Random Forest

```{r pressure, echo=FALSE}
rf_model <- randomForest(Response ~ ., data = train_data, ntree = 100)
rf_pred <- predict(rf_model, test_data, type = "prob")[, 2]
```

## Threshold Optimization

```{r pressure, echo=FALSE}
# Function to calculate profit at a given threshold
calculate_profit <- function(predictions, actual, threshold, cost_per_contact, revenue_per_response) {
  predicted_positives <- predictions >= threshold
  positives_correctly_predicted <- sum(predicted_positives & (actual == 1))
  total_contacted <- sum(predicted_positives)
  profit <- positives_correctly_predicted * revenue_per_response - total_contacted * cost_per_contact
  return(profit)
}

# Define parameters
cost_per_contact <- 3
revenue_per_response <- 11
thresholds <- seq(0.1, 0.9, length.out = 40)

# Calculate profits for each threshold
profits_log <- sapply(thresholds, calculate_profit, 
                      predictions = lognet_pred, 
                      actual = as.numeric(test_data$Response) - 1, 
                      cost_per_contact = cost_per_contact, 
                      revenue_per_response = revenue_per_response)

profits_rf <- sapply(thresholds, calculate_profit, 
                     predictions = rf_pred, 
                     actual = as.numeric(test_data$Response) - 1, 
                     cost_per_contact = cost_per_contact, 
                     revenue_per_response = revenue_per_response)

# Find optimal thresholds
optimal_threshold_log <- thresholds[which.max(profits_log)]
optimal_profit_log <- max(profits_log)

optimal_threshold_rf <- thresholds[which.max(profits_rf)]
optimal_profit_rf <- max(profits_rf)

cat("Logistic Regression: Optimal Threshold =", optimal_threshold_log, "with Profit =", optimal_profit_log, "\n")
cat("Random Forest: Optimal Threshold =", optimal_threshold_rf, "with Profit =", optimal_profit_rf, "\n")

```

## Confusion Matrices

```{r pressure, echo=FALSE}
# Final predictions based on optimal thresholds
lognet_class <- ifelse(lognet_pred >= optimal_threshold_log, 1, 0)
rf_class <- ifelse(rf_pred >= optimal_threshold_rf, 1, 0)

# Confusion matrices
confusion_matrix_log <- table(Predicted = lognet_class, Actual = as.numeric(test_data$Response) - 1)
confusion_matrix_rf <- table(Predicted = rf_class, Actual = as.numeric(test_data$Response) - 1)

cat("Confusion Matrix - Logistic Regression:\n")
print(confusion_matrix_log)

cat("\nConfusion Matrix - Random Forest:\n")
print(confusion_matrix_rf)

```

## Visualization of Profits by Threshold

```{r pressure, echo=FALSE}
plot(thresholds, profits_log, type = "l", col = "blue", lwd = 2, 
     xlab = "Threshold", ylab = "Profit", main = "Profit vs Threshold")
lines(thresholds, profits_rf, col = "red", lwd = 2)
legend("bottomright", legend = c("Logistic Regression", "Random Forest"),
       col = c("blue", "red"), lwd = 2)

```
